---
title: "Session 6"
subtitle: "Ecological Data Collection and Processing: Exercises"
author: "Ward Fonteyn & Stef Haesen"
date: "25/11/2021"
knit: (function(inputFile, encoding){
  rmarkdown::render(inputFile, encoding = encoding, output_file = "session-6-document.html", output_format = "html_document",  output_dir = "knitted");
  rmarkdown::render(inputFile, encoding = encoding, output_file = "session-6-slides.html", output_format = "ioslides_presentation", output_options = list(widescreen = T),
  output_dir = "knitted") })
---

```{r setup, include=F, echo=F}
knitr::opts_chunk$set(echo = FALSE)
library(DiagrammeR)
library(knitr)
library(kableExtra)
library(vegan)
library(ape)
library(FD)
library(edcpR)
```


# Session 6: Ordination

## Ordination

- Multivariate analysis
- Detect patterns
- Emphasis on continuity in the data

Goal: reduce dimensions

## Ordination methods

- PCA: Principal component analysis
- PCoA: Principal coordinate analysis
- NMDS: Non-metric multidimensional scaling

## Datasets

- varespec: cover values of 44 psecies in 24 plots in lichen pastures
- varechem: soil characteristics for the same plots

## Exercises

**PCA**

- Perform a PCA on the environmental data and plot the results.
- Extract the eigen values and calculate the amount of variation explained by each axis.

**PCoA**

- Repeat the same tasks with the PCoA.

**NMDS**

- Perform a NMDS on the species data. Which dissimilarity index should be used?
- Try to plot the results and add the environmental variables to the plot.

## Part 1

```{r, echo = T}
# First update package
library(edcpR)
data(varechem_s6, package = "edcpR")
```

**PCA**

- Perform a PCA on the environmental data and plot the results.
  - **TIP**: vegan package
- Extract the eigen values and calculate the amount of variation explained by each axis.

## Loading data

```{r, echo = T}

data(varechem_s6, package = "edcpR")
env <- varechem_s6
str(env[,10:17])
env$Managementtype <- as.factor(env$Managementtype)
env$Use <- as.factor(env$Use)
```

## PCA (1)
```{r, echo = T}
# rda without conditions or constraints is PCA
# We have nutrient contents (g/kg), soil depth (cm), pH (1-14)...
# scale = TRUE for data on different scales!!
PCA <- rda(env[,2:15], scale=TRUE) 
```

## PCA (2)
```{r, echo = T, fig.height = 3.5}
# Bar plot of relative eigen values (= percentage variance explained by each axis)
eig <- as.vector(PCA$CA$eig)
barplot(eig/sum(eig), xlab = "Principal Component", 
          ylab = "Explained variance", ylim = c(0, 0.5))
```

## PCA (3)
```{r, echo = T}
# Calculate the percent of variance explained by first two or three axes
# First two axes explain 60% of the variance
# First three axes explain 72% of the variance
sum((eig/sum(eig))[1:2])
sum((eig/sum(eig))[1:3])
# Three dimensional representation is possible, but more difficult
# We prefer to plot in two dimensions, but not possible here
# Solution: plot PC1 Vs. PC2 & PC1 vs. PC3
```

## PCA (4)

```{r, echo = T}
# In a biplot of a PCA, environment's scores are drawn as arrows
# These arrows point in direction of increasing value for that variable
biplot(PCA, choices=c(1,2), type=c("text","points")) 
```

## PCA (5)

```{r, echo = T}
# If we would like to include three dimensions: biplot of axis 1 vs 3
biplot(PCA, choices=c(1,3), type=c("text","points")) 
```

## Part 2

**PCoA**

- Repeat the same tasks with the PCoA
  - **TIP**: ape package
  - **TIP**: gowdis() function from FD package

## PCoA (1)

```{r, echo = T}
# First step: create a distance matrix
# We have quantitative and categorical environmental variables
# We thus use Gower distance for mixed variables
dist <- gowdis(env[, 2:17])

```

## PCoA (2)

```{r, echo = T, fig.height = 3.5}
# Perform PCoA & plot the eigenvalues
PCoA <- pcoa(dist)
barplot(PCoA$values$Relative_eig, xlab = "Principal Component", 
          ylab = "Explained variance", ylim = c(0, 0.45)) 

```

## PCoA (3)

```{r, echo = T}
# Some distance measures may result in negative eigenvalues --> add a correction
PCoA <- pcoa(dist, correction = "cailliez")
barplot(PCoA$values$Rel_corr_eig, xlab = "Principal Component", 
          ylab = "Explained variance", ylim = c(0, 0.3))

```

## PCoA (4)

```{r, echo = T, warning=F}
# There are no variables plotted on this biplot (input = dissimilarity matrix) 
biplot.pcoa(PCoA)
```

## PCoA (5)

```{r, echo = T, warning=F}
# However, we could work around this problem
biplot.pcoa(PCoA, env[6])
```

## Part 3

**NMDS**

- Perform a NMDS on the species data. Which dissimilarity index should be used?
  - **TIP**: metaMDS function from vegan package
- Try to plot the results and add the environmental vairables to the plot
  - **TIP**: envfit function from vegan package

## NMDS (1)

```{r, echo = T, eval = T}
data(varespec_s6, package = "edcpR")
sp <- varespec_s6

# Determine the number of dimensions for the final NMDS
# Run NMDS for 1 up to 10 dimensions
# Bray-Curtis distance matrix for species matrix with cover data
# If you don`t provide a dissimilarity matrix, metaMDS automatically applies Bray-Curtis
A<-vector() 
B<-vector() 
for (i in 1:10) {
  NDMS <- metaMDS(sp[2:45], distance = "bray", k=i, autotransform = FALSE, trymax=100)
  solution <- NDMS$converged
  stress <- NDMS$stress
  A1 <- cbind(solution,stress)
  B <- rbind(B,i)
  A <- rbind(A,A1)
}
```

## NMDS (2)

```{r, echo = T, eval = T}
#Combine results of all NMDS
NMDSoutput<-data.frame(B,A)
names(NMDSoutput) <- c("k","Solution?","Stress")
NMDSoutput
```

## NMDS (3)

```{r, echo = F}
plot(NMDSoutput$Stress, xlab = "# of Dimensions", ylab = "Stress", main = "NMDS stress plot")

```

## NMDS (4)

```{r, echo = F}
plot(NMDSoutput$`Solution?`, xlab = "# of Dimensions", ylab = "Solution", main = "NMDS convergence plot")
```

## NMDS (5)

```{r, echo = T, eval=T}
# Final result depends on the initial random placement of the points 
# We need set a seed to make the results reproducible
set.seed(999)
NMDSfinal <- metaMDS(sp[2:45], distance = "bray", k=2, autotransform = FALSE, trymax=100 )
```

## NMDS (6)

```{r, echo = T, eval= T}
# We can again check the stress
NMDSfinal$stress
```

## NMDS (7)

```{r, echo = T, eval= T}
# Let’s check the results of NMDSfinal with a stressplot
stressplot(NMDSfinal)
```

## NMDS (8)

```{r, echo = T}
#Plot results wit text for species
plot(NMDSfinal, type = "t")
```

## NMDS (9)

```{r, echo = T}
#Plot results wit points for species
plot(NMDSfinal, type = "p")
```

## NMDS (10)

```{r, echo = T}
# There are no environmental variables on the plot (same as PCoA)
# We can again work around this
ef <- envfit(NMDSfinal, env[2:10], permu = 999)
```

## NMDS (11)

```{r}
ef
```

## NMDS (12)

```{r, echo = T}
# Plot the vectors of the significant correlations
plot(NMDSfinal, type = "t", display = "sites")
plot(ef, p.max = 0.05)

```

## Assignment

There is NO assignment for this week's session:

 - Combined assigned for the session on ordination and classification
 - Online after the session on classification (December 9th)
 
Remember to upload Assignment 4 (Relationships & Comparing Means) by next week: **December 2nd, 12 pm (= at noon!)**
