---
title: "Assignment 5"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{assignment-4}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(edcpR)
```

Use the following datasets:

```{r, echo = T}
library(edcpR)
data(GutterTor_sp, GutterTor_env, package = "edcpR")
```

**Data**: GutterTor species (cover data) and GutterTor_env

Please complete the following problems:

- Only analyse relevant variables.
- Perform an ordination on the vegetation data.
- Classify the different plots into different clusters based on environmental characteristics.
- Visualize the ordination. Add the different clusters on this ordination plot, together with the significant environmental variables.

Remember:

- **annotate** your script using comments


Submit your R-script as **Surname_FirstName_assignment5.R**

Upload everything on Toledo before **December 8, 12 pm** (= at noon)!