---
title: "Assignment 1"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{assignment-1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(edcp)

```

Please complete the following problems.

Remember:

- R is case-sensitive. Make sure you store your answers using the exact name specified.
- You will be graded on:
  - The **correctness** of your answers.
  - The **efficiency** of your code. Don't make your script unnecessarily long and only do the essential steps.
  - The **clarity** of your code. Make sure you explain the steps in your script with annotations (`#`). Without comments we don't know if you understand what is happening.

## Data

Load in the data you will need as follows:

```{r}
data(vegdata1, vegdata2)
```

## Part 1

Extract the following values from the *vegdata1* dataset:

Make sure 

- Store the column with name *Elevation* in a vector called **my_vector**.
- Store the number in the *18th* row and *6th* column in an object called **my_number**.
- Create a new vector **my_vector_2** which contains both *my_vector* and *my_number*.
- Calculate the mean of the *12th* column and store it in **my_mean**.
- What kind of class is stored in the *sixth* column? Store this as a string in **my_data_type**.
- What are the dimensions of *vegdata1*? Store this as a vector in **my_dim**.

## Part 2

Merge *vegdata1* and *vegdata2* into one data frame. Store your result in **my_data_frame**. Hint: the resulting dataframe has 144 rows.

## Part 3

Use the package *"dplyr"* to only select rows containing data from *Grassland* from your merged data frame. Store your result in **my_grassland**. 

## Submit your work

Save your R-script as **Surname-FirstName-assignment-1.R**. Upload this on Toledo before **October 6th, 12:00 noon** in the assignments folder!
