---
title: "Assignment 2"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{assignment-2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(edcpR)
```

Please complete the following problems.

Remember:

- R is case-sensitive. Make sure you store your answers using the exact name specified.
- You will be graded on:
  - The **correctness** of your answers.
  - The **efficiency** of your code. Don't make your script unnecessarily long and only do the essential steps.
  - The **clarity** of your code. Make sure you explain the steps in your script with annotations (`#`). Without comments, we don't know if you understand what is happening.

## Data

Load in the data you will need as follows:

```{r}
data(vegdata_a2,  package = "edcpR")
```

## Part 1

Check the structure of the *vegdata_a2* dataset: 

- Correct data types if incorrect.
  - **TIP**: cover values are best converted to numerical. 
  - **TIP**: Character types which show clear categories are best converted to factor.
- Check for NA's. Remove rows containing an NA-value from the data frame.
- Check for incorrect values. Correct if necessary.
- Assign your final, clean data frame to an object called **my_df_1**.

## Part 2

Are there any outlier values in the **Elevation** variable from **my_df_1**? If so, identify the outliers and remove the corresponding rows from the data frame. Save the new data frame in an object called **my_df_2**.

## Part 3

Create a scatterplot using the ggplot2 package, meeting the following criteria:

- Use data from *my_df_2*.
- Elevation plotted on the x-axis & Litter cover on the y-axis.
- Use different colours for each Vegetation Type (i.e. Forest, Grassland & Shrubland)
- Use a theme with white background and gray grid lines
- Add a centred title: "Relation between Litter Cover and Elevation"
- Add a centred subtitle: "Relationship broken down by Vegetation Type"
- Add your name as a caption to the figure.

## Upload

Upload the following files to Toledo:

- Your R-script as **Surname_FirstName_assignment2.R**.
- Your scatterplot as **Surname_FirstName_assignment2.png**. Make sure everything on the graph is readible!

Remember to upload everything before **October 27, 12 am (= at noon!)**.
