---
title: "Assignment 3"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{assignment-3}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(edcpR)
```

Please complete the following problems.

Remember:

- R is case-sensitive. Make sure you store your answers using the exact name specified.
- You will be graded on:
  - The **correctness** of your answers.
  - The **efficiency** of your code. Don't make your script unnecessarily long and only do the essential steps.
  - The **clarity** of your code. Make sure you explain the steps in your script with annotations (`#`). Without comments, we don't know if you understand what is happening.

## Data

Load in the data you will need as follows:

```{r}
data(vegdata_a2,  package = "edcpR")
```

## Part 1

What can you tell on the species diversity in Leuven? 

- At plot level? Calculate all relevant indices. Make clear which index you calculate with annotations.
- Between plots? How many plot pairs are perfectly dissimilar? Which pair of distinct plots shows the highest similarity? Answer by means of annotations within your R-script!

## Part 2

Vegetation sampling has also been performed in the Hoge Kempen National Park (a region to the North-West of Leuven). What can you say on the species diversity from that region compared to Leuven? Answer by means of annotations within your R-script!

Remember to delete the line of code with **setwd()**. Save your R-script as **Surname_FirstName_Assignment3.R**.

Upload this document on Toledo before **October 28, 12 pm** (= at noon)!