---
title: "Session 6"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{session-6}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(edcpR)
```

## Part 1

```{r}
# First update package
library(edcpR)
data(varespec_s6, package = "edcpR")
data(varechem_s6, package = "edcpR")
```

**PCA**

- Perform a PCA on the environmental data and plot the results.
  - **TIP**: vegan package
- Extract the eigen values and calculate the amount of variation explained by each axis.

## Part 2

**PCoA**

- Repeat the same tasks with the PCoA
  - **TIP**: ape package
  - **TIP**: gowdis() function from FD package
  
## Part 3

**NMDS**

- Perform a NMDS on the species data. Which dissimilarity index should be used?
  - **TIP**: metaMDS function from vegan package
- Try to plot the results and add the environmental vairables to the plot
  - **TIP**: envfit function from vegan package
  
## Assignment

There is NO assignment for this week's session:

 - Combined assigned for the session on ordination and classification
 - Online after the session on classification (December 9th)
 
Remember to upload Assignment 4 (Relationships & Comparing Means) by next week: **December 2nd, 12 pm (= at noon!)**