---
title: "Session 2"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{session-2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(edcpR)
```


## Data

use the dataset "varechem_s2" provided in the course pacakge:

```{r}
data(varechem_s2, package= "edcpR")
```

## Tips 

Find useful tips and function suggestions on the cheat-sheet vignette:

```{r,  echo=T, eval=F}
browseVignettes(package = "edcpR")
```

## Part 1

- Check the structure and data type of the data (and change in R if necessary)
  - **Extra**: Try to assign meaningful levels for the variable *Managementtype*
- Check NA and incorrect values (and change in R if necessary)
  - **TIP**: anyNA(), complete.cases(), ...
  - Calculate summary metrics (min, max, mean, ...) to check for anomalies
- Visualize your data
  - Try different types of graphs (scatterplot, histogram, boxplot...)
  
  
## Part 2

- Perform outlier analysis
  - for a univariate case: let's consider S
  - for a multivariate case: let's consider P~K here
    - do we have influential outliers?
    - **TIP**: olsrr-package
    
## Part 3

- Check distribution of variables
  - Is the data normally distributed?
  - Make a visual inspection for the variable Potassium **(K)** and Calcium **(Ca)**
  - Use a statistical test
- Apply transformations to normalize variables
  - Let's do this for the variable Calcium **(Ca)**
- Apply transformations to linearize relationships
  - Do this for the relationship between Iron **(Fe)** and Zinc **(Zn)**
  
## Assignment 2

Access the assignment:

```{r}
vignette("assignment-2", package = "edcpR")
```

Remember to upload everything before **October 27, 12 am (= at noon!)**.

