---
title: "Session 3"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{session-3}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(edcpR)
```

## Part 1

```{r, echo=T, eval=F}
library(vegan)
data("BCI", package = "vegan")
```

- Alfa diversity
  - Calculate species richness, shannon and simpson indices
    - **TIP**: `vegan` package
  - Calculate corresponding evenness indices
    - **TIP**: use formulas provided in the lecture slides
  - Calculate corresponding Hill numbers
    - **TIP**: `tsallis()` function
  - Plot relationships between different alfa diversity measures
  
## Part 2

- Beta diversity
  - Calculate pairwise beta diversity
    - **TIP**: `betadiver()`
  - Partition beta diversity in nestedness and turnover components
    - **TIP**: `betapart` package
    - **TIP**: tranform to presence/abscence data!
  - Calculate multiple site beta diversity
    
## Part 3

- Gamma diversity
  - Calculate chao estimators for the BCI dataset 
    - **TIP**: `specpool()`
  - An old dataset has the following Chao estimators: Chao = 280 (SE = 11.5)  
    - Did we lose or gain species?
  
## Part 4

```{r, echo=T, eval=F}
library(FD)
data("tussock", package = "FD")
```

- Functional diversity
  - Calculate the functional diversity indices from the Tussock dataset:
    - **TIP**: `tussock` dataset from `FD` package
    - **TIP**: `dbFD()`
    
## Assignment 3

Access the assignment:

```{r, echo=T, eval=F}
vignette("assignment-3", package = "edcpR")
```

Remember to upload everything before **November 3rd, 12 am (= at noon!)**.


