---
title: "Session 7"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{session-7}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(edcpR)
```

# Data

The first three exercises use the `iris`dataset:

```{r, echo = T}
library(datasets)
data(iris, pacakges = "datasets")
```

The last exercise uses the varespec dataset:
```{r, echo = T}
library(vegan)
data(varespec, pacakges = "vegan")
```

## Hierarchical clustering
  
- Construct a dissimilarity matrix for the sepal length
- Perform an agglomerative hierarchical clustering of the dataset
  - based upon the sepal length
  - **TIP**: `cluster` package
- Plot the dendrogram of the performed clustering
  - change the output from the agglomerative clustering into a hclust object
- Find the optimum number of clusters
  - **TIP**: `stats` package
- How does the clustering changes when you use 4 clusters instead of 3?
- Visualize the different clusters by plotting rectangles on the dendrogram

## K means

- Find the optimum number of clusters
  - **TIP**: `cluster` package
- Perform a K means clustering of the iris dataset with selected number of clusters
  - based upon the petal length & width
- Visualize the clustering using ggplot2

## K nearest neighbours

- Split data into random training and random testing sets
  - use a random 70%-30% split
- Find the optimum number of neighbours
  - **TIP**: `class` package
- Visualize the clustering using ggplot2

## Indicator species analysis

- Use the varespec dataset
- Divide the species data in clusters
- Find indicator species for each cluster
  - **TIP**: `indicspecies` package

## Assignment 5

Access the assignment:

```{r, echo=T, eval=F}
vignette("assignment-5", package = "edcpR")
```

Submit your R-script as **Surname_FirstName_assignment5.R**

Upload everything on Toledo before **December 8, 12 pm** (= at noon)!
